---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
```

### Off-pump coronary artery bypass grafting is safe and effective in patients with severe left ventricular dysfunction

Authors: Mateo Marin-Cuartas^1^, Salil V. Deo^2^, Paulina Ramirez^1^, Alexander Verevkin^1^, Sergey Leontyev^1^, Michael A. Borger^1^, Piroze M. Davierwala^1^

<u>Affiliations:</u>


- ^1^ University Department of Cardiac Surgery, Heart Center Leipzig, Leipzig, Germany, ^2^ Louis Stokes Cleveland VA Medical Center, Department of Veterans Affairs, Cleveland Ohio, United States of America.

<u>Corresponding Author:</u>

Dr. Piroze M. Davierwala,

University Department of Cardiac Surgery,Heart Center Leipzig,Struempellstrasse 39,
04289 Leipzig, Germany

Email: pirarm@hotmail.com


```{r, echo = F, message = F, warning=F}


```



This repository contains code and script used for the analysis in the manuscript. The repo does not contain any patient information or patient identifiers.

<u>**Key question:**</u> Is off-pump coronary artery bypass (CABG) safe and effective in patients with poor left ventricular (LV) function?

<u>**Key findings:**</u> Similar mortality risk between on- and off-pump CABG (P=0.7); Incomplete revascularization more common in off-pump CABG

<u>**Take-home message:**</u> Off-pump CABG is safe and effective in patients with LV dysfunction. Target complete revascularization to improve long-term outcomes.
